High frequency intraday forex rates forecasting using an online cointegration approach

### 2015-06-23 manuscript/manuscript.tex
Primera version articulo

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
\usepackage{amsfonts,amssymb,amsmath,
            amssymb,amsopn,amsxtra,amstext,amsbsy}
\usepackage{epsfig}
\usepackage[makeroom]{cancel}
\usepackage{array,arydshln}
\usepackage{array}
\usepackage{url}
%\usepackage{tikz}
\usepackage[utf8x]{inputenc}
\usepackage{fancyhdr}
\usepackage{hyperref,wasysym}
\usepackage{color}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{multirow}
%\usepackage[spanish]{babel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{xcolor}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{pmat}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newtheorem*{myteo}{Teorema}
%\newtheorem*{mydef}{Definici\'on}
%% MYREMARK ENVIRONMENT %%%%%%%%%%%
\newenvironment{mycolor}{\color{blue}}{}
\newcounter{myremarkcount}
\setcounter{myremarkcount}{0}
\newsavebox{\myremarkbox}
\newenvironment{myremark}
{% This is the begin code
\medskip
\begin{flushright}
\begin{lrbox}{\myremarkbox}%
\begin{minipage}[t]{1.00\linewidth}
\stepcounter{myremarkcount}
\footnotesize
\begin{mycolor}
{{\bf Proofreading Remark.}\ }
% \begin{it}
}
{% This is the end code
% \end{it}
\end{mycolor}
\end{minipage}
\end{lrbox}
\fbox{\usebox{\myremarkbox}}
\end{flushright}

\bigskip}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{High frequency Intraday Forex rates forecasting using an online cointegration approach}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

VECM parameters are obtained using ordinary least squares
(OLS) method \cite{golub1980}. OLS has two main problems: it is sensitive to errors on input data
and also involves many calculations. The former problem is commonly solved using
Ridge Regression (RR) \cite{hoerl1970} which introduces a regularisation
parameter that leads to an unbiased estimation with better generalisation
capability. The second problem of computational complexity depends on the number
of past values and observations considered.  Recently, online learning
algorithms have been proposed to solve problems with large data sets because of
their simplicity and their ability to update the model when new data is
available. We use an online version of the linear ridge regression and also a modification introduced by 
Vovk \cite{vovk2001} called the Aggregating algorithm regression (AAR), which is also known as Vovk-Azoury-Warmuth predictor ~\cite{azoury2001}.

Our proposal is an online formulation of the VECM called Online VECM (OVECM)
based on the consideration of only a sliding window of the historical data at every step.  OVECM parameters are obtained via OLS, RR or AAR. In order to measure performance, a competitive analysis is done which compares the proposal cumulative loss against the optimal offline algorithm.

@@

The solution $\widehat{\mathbf{A}}$ to
equation~(\ref{eq:vareq}) can be obtained by the ordinary least squares (OLS)
method. $\widehat{\mathbf{X}}$ is the solution of the quadratic optimization problem
\begin{equation*}
\widehat{\mathbf{X}} = \underset{\mathbf{X}}{\text{Arg\;min}}
\|\mathbf{A}\mathbf{X}-\mathbf{Y}\|_2^2
\end{equation*}
\noindent for which the solution $\widehat{\mathbf{X}}$ is well-known:
\begin{equation*}
\label{eq:MP}
\widehat{\mathbf{X}}=\mathbf{A}^{\!\!+}\,\mathbf{Y}
\end{equation*}
\noindent where $\mathbf{A}^{\!\!+}$ is the Moore-Penrose pseudo-inverse
which, when $\mathbf{A}$ is full rank, can be written as follows: 
\begin{equation}
\label{eq:pseudoinverse}
\mathbf{A}^{\!\!+}= (\mathbf{A}^{\!\!\top} \mathbf{A})^{-1}\mathbf{A}^{\!\!\top} \, .
\end{equation}
%when $\mathbf{A}$ is not full rank, i.e
%$rank(\mathbf{A})=k <  n \leq m$, $\mathbf{A}^\top \mathbf{A}$ is
%always singular and equation~(\ref{eq:pseudoinverse}) cannot be used.
%More generally, the pseudo-inverse is best computed using the compact
%singular value decomposition (SVD) of $\mathbf{A}$ which is:
%\begin{equation*}
%    \label{eq:compactsvd}
%    \mathbf{A}=
%    \mathbf{U_1}
%    \boldsymbol \Sigma_1
%    \mathbf{V}_1^{\top} \, .
%\end{equation*}
%Pseudo-inverse can then be written as follows:
%\begin{equation*}
%\label{eq:pseudoinversesvd}
%\mathbf{A}^{\!\!+} = \mathbf{V}_1 \boldsymbol \Sigma_1^{-1} \mathbf{U}_1^\top \, .
%\end{equation*}

@@

In batch learning, the training phase is commonly a computationally expensive
process. Therefore, when new data arrives it can't be included easily to the
model. Moreover, it could happen that there is  won't have enough time to process the
new data before more data arrives. In online learning algorithms there is no
training phase, but the model is updated and evaluated at every time step. This
model updating is computationally less expensive than a training phase.
Online algorithms allow incremental learning by processing one instance at a
time. This is done updating the current model instead of building the model from
scratch.

@@

\noindent where $l$ is some loss function. Performance is later measured after
$T$ trials as:

@@

Competitive analysis was designed for analysing the performance of an online algorithm compared with its optimal offline algorithm. An optimal offline algorithm can view the sequences of requests in advance. The effectiveness of an online algorithm \cite{sleator1985} may be measured by its competitive ratio which is the worst-case ratio of the online algorithm and the the optimal offline algorithm.

@@

Online RR is the online formulation of the regularised Least Squares method
and is based on the following equivalent formulation of the RR optimal solution.

%Since 
%\begin{equation}
%\label{eq:notation}
%	\mathbf{A} = 
%\left[
%  \begin{tabular}{c>{$}c<{$}c}
%    --- & \mathbf{a}^{\top}_{1} & ---\\
%    --- & \mathbf{a}^{\top}_{2} & ---\\
%    & \vdots & \\
%    --- & \mathbf{a}^{\top}_{m} & ---
%  \end{tabular}
%\right]
%\quad \text{and} \quad
%\mathbf{B} =
%\left[
%  \begin{tabular}{c>{$}c<{$}c}
%    --- & \mathbf{b}_{1} & ---\\
%    --- & \mathbf{b}_{2} & ---\\
%    & \vdots & \\
%    --- & \mathbf{b}_{m} & ---
%  \end{tabular}
%\right]
%\end{equation}

@@

@@

The AAR, proposed by~\cite{vovk2001} (also known as Vovk-Azoury-Warmuth predictor ~\cite{azoury2001}), is an application of the aggregating
algorithm to the problem of regression. The idea is introduce the new input
vector $\mathbf{x}_{m+1}$ to solve the model parameters:

\begin{equation}
\label{eq:AARexpand}
\mathbf{X}_{aar} = \displaystyle \big (\sum_{t=1}^{m+1}
\mathbf{a}_t \mathbf{a}_t  ^\intercal + \gamma \mathbb{I}\big )^{-1}
\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t \, .
\end{equation}

If we define $\displaystyle\mathbf{S}= \sum_{t=1}^{m+1} \mathbf{a}_t
\mathbf{a}_t  ^\intercal + \gamma \mathbb{I} $ and $\mathbf{W}=
\displaystyle\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t$, the
algorithm~\ref{alg:AAR} is slightly different to the algorithm~\ref{alg:RR}, 
which updated matrix $\mathbf{S}$ before making the prediction:

@@

Since financial problems are stream data problems, it is unfeasible to include
all input data in a VECM model. Our proposal consists on an online version of
VECM (OVECM) capable of updating the model with new arrival data and give
responses in a short period of time. OVECM only considers a time varying window
of historical data and the using of RR or its variant AAR (Aggregating Algorithm
for Regression) as an alternative of OLS to get model parameters.

On the other hand, getting cointegration vectors using Johansen method is also
an expensive procedure. However, since cointegration
vectors represent the long-run relationship between the time series, they
vary little in time. Our proposal obtains new cointegration vectors only when
this long-term relationship changes. This change is detected by tracking 
the Mean Absolute Percentage Error (MAPE) of the last $n$ in-sample forecasts.

%Since VECM is a model based on time series differences, the MAPE is obtained
%from $\Delta \mathbf{y}$ as following:
% 
%\begin{equation}\label{eq:MAPE}
%\text{MAPE}[t] = \frac{1}{n} \sum_{i=1}^{n} \left| 
%\frac{\Delta \mathbf{y}_{\text{true}}[t-i]-\Delta
%\mathbf{y}_{\text{pred}}[t-i]}{\Delta \mathbf{y}_{\text{true}[t-i]}}
%\right| \, , 
%\end{equation}
%
%\noindent where $\Delta \mathbf{y}_{\text{true}}$ is the actual value of the
%time series $\mathbf{y}$ differences and $\Delta \mathbf{y}_{\text{pred}}$ is
%their forecast value.

%The method proposed consists of
%a modification of RR considering a sliding window which contains only
%the last $L$ samples %and the new input $\mathbf{x}_t$, 
%i.e. $\{\mathbf{x}_i\}_{i=t-L+1}^{t-1}$.

VECM($p$) for $l$ cointegrated variables has the following form:

\begin{align}
\label{eq:vecfull}
\Delta\mathbf{y}_t 
= \boldsymbol{\alpha\beta}^\top\mathbf{y}_{t-1} 
  + \sum_{i=1}^{p-1}\boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t  \nonumber \\
t=p+1,\dots,N,
\end{align}

Transposing each equation of the system (\ref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:vareq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}
%
\noindent where $\mathbf{B}$ dimension is $((N-p)\times l)$, $\mathbf{A}$
dimension is $((N-p)\times(r+(p-1)l +1))$, $\mathbf{X}$ dimension is $((r+(p-1)l
+1)\times l)$, $\mathbf{E}$ dimension is $((N-p)\times l)$ and $r$ is the number
of cointegration vectors:
%
\begin{alignat}{3}
\mathbf{B}
&= \begin{bmatrix}
   \Delta\mathbf{y}_{p+1}^\top \\
   \Delta\mathbf{y}_{p+2}^\top \\
   \vdots \\
   \Delta\mathbf{y}_N^\top
   \end{bmatrix}
&\quad
\mathbf{X}
&= \begin{bmatrix}
   \boldsymbol{\alpha}^\top \\
   \boldsymbol{\Phi}_1^{*\top} \\
   \boldsymbol{\Phi}_2^{*\top} \\
   \vdots \\
   \boldsymbol{\Phi}_{p-1}^{*\top} \\
   \mathbf{c}^\top
   \end{bmatrix}
&\quad
\mathbf{E}
&= \begin{bmatrix}
   \boldsymbol{\epsilon}_{p+1}^\top \\
   \boldsymbol{\epsilon}_{p+2}^\top \\
   \vdots \\
   \boldsymbol{\epsilon}_N^\top \\
   \end{bmatrix}
\end{alignat}
\noindent and 
\begin{align}
\mathbf{A} 
&= \begin{pmat}[{....|}]
   \mathbf{y}_p^\top \boldsymbol{\beta} & \Delta \mathbf{y}_p^\top & \Delta\mathbf{y}_{p-1}^\top & \dots 
                    & \Delta\mathbf{y}_2^\top & 1 \cr
   \mathbf{y}_{p+1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{p+1}^\top & \Delta\mathbf{y}_p^\top & \dots
                       & \Delta\mathbf{y}_3^\top & 1 \cr
   \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \cr
   \mathbf{y}_{N-1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{N-1}^\top & \Delta\mathbf{y}_{N-2}^\top & \dots 
                       & \Delta\mathbf{y}_{N-p-1}^\top & 1 \cr
   \end{pmat}\, .
\end{align}
%Taking into account the error term $\mathbf{E}$, equation~(\ref{eq:vareq}) 
%can be solved with respect to $\mathbf{X}$ using the ordinary least
%squares estimation.
However, the number of rows of matrix $\mathbf{A}$ increases with the number of
observations $\mathbf{y}_t$. Our proposal considers only the most recent $L$
rows of matrices $\mathbf{A}$ and $\mathbf{B}$ defined as  $\mathbf{A}(t)$ and
$\mathbf{B}(t)$:

\begin{equation}
\label{eq:notation}
	\mathbf{A}(t) = 
\left[
  \begin{tabular}{c>{$}c<{$}c}
    --- & \mathbf{a}^{\top}_{t-L} & ---\\
    --- & \mathbf{a}^{\top}_{t-L+1} & ---\\
    & \vdots & \\
    --- & \mathbf{a}^{\top}_{t} & ---
  \end{tabular}
\right]
\quad \text{and} \quad
\mathbf{B}(t) =
\left[
  \begin{tabular}{c>{$}c<{$}c}
    --- & \mathbf{b}_{t-L} & ---\\
    --- & \mathbf{b}_{t-L+1} & ---\\
    & \vdots & \\
    --- & \mathbf{b}_{t} & ---
  \end{tabular}
\right] \, ,
\
\end{equation}

\noindent so that VECM parameters $\mathbf{X}(t)$ are found using the following
equation:

\begin{equation}
\mathbf{B}(t) = \mathbf{A}(t) \mathbf{X}(t) + \mathbf{E}(t)\\
\end{equation}

The RR solution $\mathbf{\hat{X}}(t)$ using the sliding window matrices defined above
is:

\begin{equation}
\label{eq:oproblem}
\mathbf{\hat{X}}(t)=\mathbf{S}(t)^{-1} \mathbf{W}(t) \, ,
\end{equation}

@@

\begin{eqnarray*}
\mathbf{S}(t) =&{\bf A}(t)^\top{\bf A}(t)+ \lambda \mathbb{I} &=
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{a}_{t-i}^\top + \lambda \mathbb{I}
\label{eq:S} \\
\mathbf{W}(t) =& {\bf A}(t)^\top{\bf B}(t) &= 
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{b}_{t-i} 
\label{eq:W}
\end{eqnarray*}

It is worth noticing that, at the next time step, the matrices $\mathbf{S}(t+1)$
and $\mathbf{W}(t+1)$ are slightly different to $\mathbf{S}(t)$ and
$\mathbf{W}(t)$:

\begin{eqnarray*}
\mathbf{S}(t+1)&=&
\mathbf{S}(t) +
\mathbf{a}_{t+1}
\mathbf{a}_{t+1}^\top -
\mathbf{a}_{t-L} \mathbf{a}_{t-L}^\top \\
\mathbf{W}(t+1)&=&
\mathbf{W}(t) +
\mathbf{a}_{t+1}
\mathbf{b}_{t+1} -
\mathbf{a}_{t-L} \mathbf{b}_{t-L} \, .
\end{eqnarray*}

%Agregar solo si se decide agregar Sherman-Morrison
%Therefore the Sherman-Morrison-Woodbury formula can be used to reduce inverse matrix
%calculations.

The algorithm~\ref{alg:proposal} shows OVECM using three different
methods for geting model parameters: OLS, RR, AAR which leads to three different
%algorithms OVECM-OLS, OVECM-RR, OVECM-AAR:

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$p$: number of past values \\
method: \{'OLS','RR','AAR'\} \\
$\lambda$: regularization parameter (only requires for RR and AAR methods) \\
$L$: window size of data($L<N$) \\
$\text{mean\_error}$: MAPE threshold \\
$n$: windows size to obtain in-sample MAPE \\
\ENSURE  $\,$ \\
$\{\mathbf{y}_{\text{pred}}[L+1],\dots, \mathbf{y}_{\text{pred}}[N]\}$: model predictions 
\STATE solver = new \texttt{Solver}(method,$\lambda$) \\
\FOR { $t =1$ to $N-L$ }
    \STATE $\mathbf{Y} \gets \mathbf{y}[t:t+L-1]$
    \STATE $\mathbf{y}_t \gets \mathbf{y}[t+L]$
	\IF {$t = 1$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
        \texttt{vecMatrix}(\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
    \ENDIF
	\STATE{[$\mathbf{A} \quad \mathbf{B} \quad \mathbf{a}_t \quad \mathbf{a}_L \quad \mathbf{b}_t \quad
    \mathbf{b}_L ] \gets
    \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
    \STATE $[\mathbf{X} \quad \mathbf{y}_{\text{pred}}[t]] \gets \texttt{solver.regression}
    (\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}_t)$
    \STATE $\mathbf{Y}_{\text{pred}} = \mathbf{AX}[-n:]+\mathbf{Y}[-n-1:-1]$
    \STATE $\mathbf{e} = \texttt{mape}(\mathbf{Y}_{\text{pred}},\mathbf{Y}[-n:])$
    \IF {$\texttt{mean}(\mathbf{e}) > \text{mean\_error}$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$\mathbf{A} \gets
        \texttt{vecMatrixUpdate}(\mathbf{A},\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
	    \STATE{$[\mathbf{a}_t \quad \mathbf{b}_t] \gets
        \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
        \STATE $[\mathbf{X} \quad \mathbf{y}_{\text{pred}}[t]] \gets \texttt{solver.regression}
        (\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}_t)$
    \ENDIF
\STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
\texttt{matrixUpdate}(\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}t)$}
\ENDFOR
\end{algorithmic}
\caption{OVECM: Online VECM}
\label{alg:proposal}
\end{algorithm}

@@

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}(t),\mathbf{B}(t)$ shown in equation~(\ref{eq:notation}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods.
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters $\mathbf{X}$ is also updated later.
%\item The number of cointegration vectors is set as a parameter (aun no lo
%coloco en el algoritmo).
%\item In order to set VECM parameter: $L$ and $p$, we use the Akaike Information
%Criterion (AIC). RR parameter $\lambda$ was done by cross-validation.
%\item Obtain VECM parameter $\mathbf{X}(t)$ in equation \ref{eq:optsolSLAAR} requires
%to calculate the inverse of matrix $\mathbf{S}(t+1)$ which is an expensive
%routine.However, since we already know $\mathbf{S}(t)^{-1}$ we can use
%Sherman-Morrison-Woodbury twice~\ref{eq:SMW}.
\end{itemize}

>>>>

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
\smartqed  % flush right qed marks, e.g. at end of proof
\usepackage{graphicx}
\usepackage{amsfonts,amssymb,amsmath,
            amssymb,amsopn,amsxtra,amstext,amsbsy}
\usepackage{epsfig}
\usepackage[makeroom]{cancel}
\usepackage{array,arydshln}
\usepackage{array}
\usepackage{url}
%\usepackage{tikz}
\usepackage[utf8x]{inputenc}
\usepackage{fancyhdr}
\usepackage{hyperref,wasysym}
\usepackage{color}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{multirow}
%\usepackage[spanish]{babel}

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{pmat}

@@

\title{High frequency intraday forex rates forecasting using an online cointegration approach}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

VECM parameters are obtained using ordinary least squares
(OLS) method \cite{golub1980}. OLS has two main problems: it is sensitive to errors on input data
and also involves many calculations. The former problem is commonly solved using
Ridge Regression (RR) \cite{hoerl1970} which introduces a regularisation
parameter that leads to an unbiased estimation with better generalisation
capability. The second problem of computational complexity depends on the number
of past values and observations considered.  Recently, online learning
algorithms have been proposed to solve problems with large data sets because of
their simplicity and their ability to update the model when new data is
available. We use an online version of the linear ridge regression and also a modification introduced by 
Vovk \cite{vovk2001} called the Aggregating algorithm regression (AAR).
Our proposal is an online formulation of the VECM called Online VECM (OVECM)
based on the consideration of only a sliding window of the historical data at every step.  OVECM parameters are obtained via OLS, RR or AAR. In order to measure performance, a competitive analysis is done which compares the proposal cumulative loss against the optimal offline algorithm.

@@

The solution $\widehat{\mathbf{A}}$ to
equation~(\ref{eq:vareq}) can be obtained by the ordinary least squares (OLS)
method. $\widehat{\mathbf{X}}$ is the solution of the quadratic optimization problem
\begin{equation*}
\widehat{\mathbf{X}} = \underset{\mathbf{X}}{\text{Arg\;min}}
\|\mathbf{A}\mathbf{X}-\mathbf{Y}\|_2^2
\end{equation*}
\noindent for which the solution $\widehat{\mathbf{X}}$ is well-known:
\begin{equation*}
\label{eq:MP}
\widehat{\mathbf{X}}=\mathbf{A}^{\!\!+}\,\mathbf{Y}
\end{equation*}
\noindent where $\mathbf{A}^{\!\!+}$ is the Moore-Penrose pseudo-inverse
which, when $\mathbf{A}$ is full rank, can be written as follows: 
\begin{equation}
\label{eq:pseudoinverse}
\mathbf{A}^{\!\!+}= (\mathbf{A}^{\!\!\top} \mathbf{A})^{-1}\mathbf{A}^{\!\!\top} \, .
\end{equation}

@@

In batch learning, the training phase is commonly a computationally expensive
process. Therefore, when new data arrives it can't be included easily into the
model. Moreover, it could happen that it won't be enough time to process the
new data before more data arrives. In online learning algorithms there is no
training phase, but the model is updated and evaluated at every time step. This
model updating is computationally less expensive than a training phase.
Online algorithms allow incremental learning by processing one instance at a
time. This is done updating the current model instead of building the model from
scratch.

@@

\noindent where $l$ is some loss function. Performance is later measured after
$T$ trials using the cumulative loss function:

@@

This cumulative loss function is compared against the cumulative loss function of its its optimal offline algorithm. An optimal offline algorithm can view the sequences of requests in advance. 
This comparison is called competitive analysis  \cite{sleator1985}  which was designed for analysing the performance of an online algorithm. 
The effectiveness of an online algorithm may be measured by its competitive ratio which is the worst-case ratio of the online algorithm and the the optimal offline algorithm.

@@

Online RR is the online formulation of the regularised OLS method
and is based on the following equivalent formulation of the RR optimal solution.

@@

\noindent where 
\begin{equation}
\label{eq:notation}
	\mathbf{A} = 
\left[
  \begin{tabular}{c>{$}c<{$}c}
    -- & \mathbf{a}^{\top}_{1} & --\\
    -- & \mathbf{a}^{\top}_{2} & --\\
    & \vdots & \\
    -- & \mathbf{a}^{\top}_{m} & --
  \end{tabular}
\right]
\quad \text{and} \quad
\mathbf{B} =
\left[
  \begin{tabular}{c>{$}c<{$}c}
    -- & \mathbf{b}_{1} & --\\
    -- & \mathbf{b}_{2} & --\\
    & \vdots & \\
    -- & \mathbf{b}_{m} & --
  \end{tabular}
\right]
\end{equation}

@@

The AAR, proposed by~\cite{vovk2001} (also known as Vovk-Azoury-Warmuth predictor ~\cite{azoury2001}), is an application of the aggregating
algorithm to the problem of regression. The idea is to introduce the new input
vector $\mathbf{x}_{m+1}$ to solve the model parameters:

\begin{equation}
\label{eq:AARexpand}
\mathbf{X}_{aar} = \displaystyle \big (\sum_{t=1}^{m+1}
\mathbf{a}_t \mathbf{a}_t  ^\top+ \gamma \mathbb{I}\big )^{-1}
\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t \, .
\end{equation}

If we define $\displaystyle\mathbf{S}= \sum_{t=1}^{m+1} \mathbf{a}_t
\mathbf{a}_t  ^\top + \gamma \mathbb{I} $ and $\mathbf{W}=
\displaystyle\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t$, the
algorithm~\ref{alg:AAR} is slightly different to the algorithm~\ref{alg:RR}, 
which updated matrix $\mathbf{S}$ before making the prediction:

@@

Since financial problems are stream data problems, it is unfeasible to include
all input data in a VECM model. Our proposal consists on an online version of
VECM (OVECM) capable of updating the model with new arrival data and give
responses in a shorter period of time. OVECM only considers a time varying window
of historical data and the using of RR or its variant AAR (Aggregating Algorithm
for Regression) as an alternative of OLS to get model parameters.

Our proposal considers only the most recent $L$
rows of matrices $\mathbf{A}$ and $\mathbf{B}$ defined as  $\mathbf{A}_t$ and
$\mathbf{B}_t$:

\begin{equation}
\label{eq:notation}
	\mathbf{A}_t = 
\left[
  \begin{tabular}{c>{$}c<{$}c}
    -- & \mathbf{a}^{\top}_{t-L} & --\\
    %-- & \mathbf{a}^{\top}_{t-L+1} & --\\
    & \vdots & \\
    -- & \mathbf{a}^{\top}_{t} & --
  \end{tabular}
\right]
,
\mathbf{B}_t=
\left[
  \begin{tabular}{c>{$}c<{$}c}
    -- & \mathbf{b}_{t-L} & --\\
    %-- & \mathbf{b}_{t-L+1} & --\\
    & \vdots & \\
    -- & \mathbf{b}_{t} & --
  \end{tabular}
\right] \, ,
\
\end{equation}

\noindent so that VECM parameters $\mathbf{X}_t$ are found using the following
equation:

\begin{equation}
\mathbf{Y}_t = \mathbf{A}_t \mathbf{X}_t + \mathbf{E}_t\\
\end{equation}

The RR solution $\mathbf{\hat{X}}_t(\lambda)$ using the sliding window matrices defined above
is:

\begin{equation}
\label{eq:oproblem}
\mathbf{\hat{X}}_t(\lambda)=\mathbf{S}_t^{-1} \mathbf{W}(t) \, ,
\end{equation}

@@

\begin{eqnarray*}
\mathbf{S}_t =&{\bf A}_t^\top{\bf A}_t+ \lambda \mathbb{I} &=
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{a}_{t-i}^\top + \lambda \mathbb{I}
\label{eq:S} \\
\mathbf{W}_t =& {\bf A}_t^\top{\bf B}_t &= 
\sum_{i=0}^L \mathbf{a}_{t-i}\mathbf{b}_{t-i} 
\label{eq:W}
\end{eqnarray*}

It is worth noticing that, at the next time step, the matrices $\mathbf{S}_{t+1}$
and $\mathbf{W}_{t+1}$ are slightly different to $\mathbf{S}_t$ and
$\mathbf{W}_t$:

\begin{eqnarray*}
\mathbf{S}_{t+1}&=&
\mathbf{S}_t +
\mathbf{a}_{t+1}
\mathbf{a}_{t+1}^\top -
\mathbf{a}_{t-L} \mathbf{a}_{t-L}^\top \\
\mathbf{W}_{t+1}&=&
\mathbf{W}_t +
\mathbf{a}_{t+1}
\mathbf{b}_{t+1} -
\mathbf{a}_{t-L} \mathbf{b}_{t-L} \, .
\end{eqnarray*}

Since we need to obtain the inverse of $\mathbf{S}_{t+1}$ at every step,
we used the Sherman-Morrison-Woodbury formula in order to reduce the number of operations. 
The Sherman-Morrison-Woodbury formula states if a matrix $\mathbf{A}$ is a positive
definite matrix and its inverse matrix is known, then the inverse of the
matrix $\mathbf{B}=\mathbf{A} + \mathbf{x}\mathbf{x}^\top$ can be
obtained as:

\begin{equation}
\label{eq:SMW}
\mathbf{B}^{-1} = \mathbf{A}^{-1} -
\frac{(\mathbf{A}^{-1}\mathbf{x})(\mathbf{A}^{-1}\mathbf{x})^{\top}}
{1 + \mathbf{x}^{\top} \mathbf{A}^{-1} \mathbf{x}} \, .
\end{equation}

This formula reduces the inverse matrix computation order from $O(n^3)$ to
$O(n^2)$. In our case the inverse of $\mathbf{S}_{t+1}$ can be obtained applying twice the Sherman-Morrison-Woodbury formula.

The algorithm~\ref{alg:proposal} shows OVECM using three different
methods for geting model parameters: OLS, RR, AAR which leads to three different
settings for OVECM:

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$p$: number of past values \\
method: \{'OLS','RR','AAR'\} \\
$\lambda$: regularization parameter (only requires for RR and AAR methods) \\
$L$: window size of data($L<N$) \\
$\text{mean\_error}$: MAPE threshold \\
$n$: windows size to obtain in-sample MAPE \\
\ENSURE  $\,$ \\
$\{\mathbf{y}_{\text{pred}}[L+1],\dots, \mathbf{y}_{\text{pred}}[N]\}$: model predictions 
\STATE solver = new \texttt{Solver}(method,$\lambda$) \\
\FOR { $t =1$ to $N-L$ }
    \STATE $\mathbf{Y} \gets \mathbf{y}[t:t+L-1]$
    \STATE $\mathbf{y}_t \gets \mathbf{y}[t+L]$
	\IF {$t = 1$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$\mathbf{A} ,\quad \mathbf{B} \gets
        \texttt{vecMatrix}(\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
    \ENDIF
	\STATE{ $\mathbf{a}_t ,\quad \mathbf{b}_t \gets
    \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
    \STATE $\mathbf{X} ,\quad \mathbf{Y}_{\text{pred}}[t] \gets \texttt{solver.regression}
    (\mathbf{a}_t,\mathbf{b}_t)$
 	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
\ENDFOR
\end{algorithmic}
\caption{OVECM: Online VECM}
\label{alg:proposal}
\end{algorithm}

@@

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. The first cointegration vector is only considered.
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}_t,\mathbf{B}_t$ shown in equation~(\ref{eq:notation}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods. This solver class obtains the inverse of matrix $\mathbf{S}_{t+1}$ using the  Sherman-Morrison-Woodbury twice (see equation \ref{eq:SMW}).
\item In order to set VECM parameter: $L$ and $p$, we use the Akaike Information
Criterion (AIC). RR parameter $\lambda$ was done by cross-validation.
\end{itemize}
